---
title: "SARS-CoV-2 reactive T-cells in COVID-19 - Figure 5"
output:
  html_document:
    toc: yes
    theme: united
    code_folding: hide
  pdf_document:
    toc: yes
---

```{r, echo = F}
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "", fig.width = 9, fig.height = 6)
#knitr::opts_chunk$set(dpi=150,fig.width=7)
```

```{r load_libraries, echo = F, message = F, warning = F}
library(sjPlot)
library(sjmisc)
library(readxl)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(broom)
library(stringr)
library(MASS) # for rlm
library(ggthemes)
library(lmerTest)
library(glmm)
library(ggstatsplot)
#library(patchwork)
#library(kableExtra)
#library(rmarkdown)
```

## Load data tables
```{r load_datasets}
# delta alpha clinical table
PNT <- read_excel("../2022_05_19_PNT_CT.xlsx") %>% 
  dplyr::select(contains("Studie 1"),
         contains("PatID"),
         contains("exclude"),
         contains("Tag PSO"),
         contains("PSO/PI"),
         contains("Zeitraum"),
         contains("Serum"),
         contains("Dexa")) %>%
  rename(PatID_Studie1 = `PatID`,
         Studie1 = `Studie 1`,
         exclude_immuno = `# exclude (immunosuppression, other)`,
         exclude_B.1.1.7 = `#exclude for B1.1.7`,
         exclude_B.1.617.2 = `#exclude for B.1.617.2`,
         TagPSO = `Tag PSO`,
         Timepoint = `Zeitraum`,
         B.1.1.7 = `Serum ID50: Variant B.1.1.7`, 
         B.1.617.2 = `Serum ID50: Variant B.1.617.2`,
         BA.2 = `Serum ID50: Variant BA.2`,
         Dexa = `Dexamethason`) %>%
  mutate(PatID_Studie1 = str_replace(PatID_Studie1, c("C19-CB-"), "")) %>% 
  mutate(PatID_Studie1 = as.character(as.numeric(PatID_Studie1))) %>% 
  mutate(TagPSO = as.numeric(TagPSO)) %>%
  mutate(Studie1 = str_replace(Studie1, c("Pa-COVID-19"), "Pa-COVID")) %>% 
  mutate(PatID_Studie1 = paste(tolower(paste(substr(Studie1, 1, 2))), 
                               tolower(paste(substr(Studie1, (nchar(Studie1)-1), nchar(Studie1)), 
                                             PatID_Studie1, 
                                             sep = "")),
                               sep = "")) %>%
  mutate(exclude_immuno = as.numeric(replace_na(exclude_immuno, 0))) %>% 
  filter(exclude_immuno == 0) %>% 
  mutate(exclude_B.1.1.7 = as.numeric(replace_na(exclude_B.1.1.7, 0))) %>% 
  #mutate(Timepoint = str_replace_all(Timepoint, c("5" = "6", "W4" = "M1"))) %>%
  mutate(exclude_B.1.617.2 = as.numeric(replace_na(exclude_B.1.617.2, 0))) %>% 
  mutate(Dexa = str_replace_all(Dexa, c("1" = "D+", "0" = "D-", "3" = "mild")))

# FACS results
facs <- read_excel("../cleaned_FACS_mastertable 2022_04.xlsx") %>% 
  rename(ID = `Studie und PatID`,
         Timepoint = `Zeitraum`) %>% 
  separate(ID, c("Studie1", "PatID_Studie1"), sep = " ") %>% 
  filter(!is.na(CEFX) | !is.na(S1) | !is.na(S2) | !is.na(SEB)) %>%
  mutate(PatID_Studie1 = as.character(PatID_Studie1)) %>% 
  #filter(!(ausschliessen == 1)) %>% 
  mutate(Studie1 = str_replace(Studie1, c("Pa-COVID-19"), "Pa-COVID")) %>% 
  mutate(PatID_Studie1 = paste(tolower(paste(substr(Studie1, 1, 2))), 
                               tolower(paste(substr(Studie1, (nchar(Studie1)-1), nchar(Studie1)), 
                                             PatID_Studie1, 
                                             sep = "")),
                               sep = "")) %>% 
  mutate(Timepoint = str_replace_all(Timepoint, c("W 2"="W2"))) %>% 
  mutate(Dexa = str_replace_all(Dexa, c("1" = "D+", "0" = "D-", "3" = "mild")))


# meta data
meta <- read_excel("../metadata 2022 05 24 CT.xlsx") %>%
  dplyr::select(#contains("sanity"),
         contains("Studie 1"),
         contains("Alter / Geschaetztes Alter bei Einschluss gr 3 Jahre"),
         contains("Geschlecht bei der Geburt"),
         contains("Dexamethason"),
         contains("Immunosuppression"),
         contains("Bemerkung"),
         contains("Start Dexa"),
         contains("Ende Dexa")) %>% 
  rename(PatID_Studie1 = `PatID Studie 1`, 
         Age = `Alter / Geschaetztes Alter bei Einschluss gr 3 Jahre`, 
         Sex = `Geschlecht bei der Geburt` ,
         Studie1 = `Studie 1`,
         Dexa = Dexamethason) %>%
  mutate(PatID_Studie1 = as.character(PatID_Studie1)) %>% 
  filter(!grepl(.$Bemerkung, pattern = "Immun", ignore.case = T)) %>% 
  mutate(PatID_Studie1 = str_replace(PatID_Studie1, c("C19-CB-"), "")) %>%
  mutate(PatID_Studie1 = as.character(as.numeric(PatID_Studie1))) %>% 
  mutate(Studie1 = str_replace(Studie1, c("Pa-COVID-19"), "Pa-COVID")) %>% 
  mutate(PatID_Studie1 = paste(tolower(paste(substr(Studie1, 1, 2))), 
                               tolower(paste(substr(Studie1, (nchar(Studie1)-1), nchar(Studie1)), 
                                             PatID_Studie1, 
                                             sep = "")),
                               sep = "")) %>% 
  mutate(Dexa = str_replace_all(Dexa, c("1" = "D+", "0" = "D-", "3" = "mild")))

# serology
sero <- read_excel("../2022_05_18_Serol_CT.xlsx") %>% 
  dplyr::select(contains("Studie 1"),
         contains("Entnahmedatum"),
         contains("Tag PSO"),
         contains("Bemerkung"),
         contains("Zeitraum"),
         contains("IgG"),
         contains("Dexa"),
         contains("ausschliessen (Immunsuppression, other)"),
         "PatID") %>% 
  rename(PatID_Studie1 = `PatID`,
         Studie1 = `Studie 1`,
         IgG = `IgG Value <7 set to 1`,
         TagPSO = `Tag PSO`, 
         Timepoint = Zeitraum, 
         Dexa = `Dexamethason`,
         ausschliessen_imm = `ausschliessen (Immunsuppression, other)`) %>% 
  mutate(ausschliessen_imm = replace_na(ausschliessen_imm, 0)) %>%
  mutate(PatID_Studie1 = str_replace(PatID_Studie1, c("C19-CB-"), "")) %>% 
  mutate(PatID_Studie1 = as.character(as.numeric(PatID_Studie1))) %>% 
  mutate(IgG = as.numeric(str_replace(IgG, c(","), "\\."))) %>% 
  filter(!ausschliessen_imm == 1) %>% 
  mutate(Studie1 = str_replace(Studie1, c("Pa-COVID-19"), "Pa-COVID")) %>% 
  mutate(PatID_Studie1 = str_replace(PatID_Studie1, c("C19-CB-"), "")) %>%
  mutate(PatID_Studie1 = as.character(as.numeric(PatID_Studie1))) %>% 
  mutate(PatID_Studie1 = paste(tolower(paste(substr(Studie1, 1, 2))), 
                               tolower(paste(substr(Studie1, (nchar(Studie1)-1), nchar(Studie1)), 
                                             PatID_Studie1, 
                                             sep = "")),
                               sep = "")) %>% 
  mutate(Timepoint = str_replace_all(Timepoint, c("5" = "6"))) %>%
  mutate(Dexa = str_replace_all(Dexa, c("1" = "D+", "0" = "D-", "3" = "mild")))

SRT <- c("S1", "S2")
strain <- c("B.1.1.7", "B.1.617.2")
level_order <- c('1', '0', '3')
```

```{r models,echo = F}

meta_age_sex <- meta %>% 
  dplyr::select(PatID_Studie1, Age, Sex)

B117 <- PNT %>% 
  filter(exclude_B.1.1.7 == 0 & Timepoint != "PI") %>% 
  dplyr::select(PatID_Studie1, Dexa, B.1.1.7, Timepoint, TagPSO) %>% 
  left_join(., meta_age_sex) %>% 
  mutate(var_col = B.1.1.7)

B16172 <- PNT %>% 
  filter(exclude_B.1.617.2 == 0 & Timepoint != "PI") %>% 
  dplyr::select(PatID_Studie1, Dexa, B.1.617.2, Timepoint, TagPSO) %>% 
  left_join(., meta_age_sex) %>% 
  mutate(var_col = B.1.617.2)

sero_m <- sero %>% dplyr::left_join(meta_age_sex) %>% 
  filter(Timepoint != "PI")

srt_m <- facs %>% 
  left_join(meta_age_sex)
```

## Time series
### SRT time series
(No patient had more than one measurement in a single time point)
```{r Timeseries T-cells, PNT and IgG of patients with multiple measurements}
## SRT discrete time points
## CD4
for(i in SRT)
  {
    print(
      facs %>%
        #filter(Marker1 == "CD4") %>%
        filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
        filter(Marker == "CD4 CD40L+ CD137+" )#| Marker ==  "CD4 CD40L+ TNF-a+" | Marker == "CD4 CD40L+ IL-21+") 
      %>% 
        #left_join(SRT_2_tp, ., by = "PatID_Studie1") %>% 
      ggplot(., aes(x = factor(Timepoint, 
                               levels = c('W2', 'M1', 'M3', 
                                          'M6')), 
                    y = !!sym(i), 
                    colour = Dexa)) +
        #geom_jitter(width = 0.1, alpha = 0.5) +
        geom_point(position = position_dodge(0.4), alpha = 0.7) +
        scale_y_log10(minor = c(seq(0.001, 0.010, by = 0.001),
                              seq(0.010, 0.100, by = 0.01),
                              seq(0.100, 1.000, by = 0.1),
                              seq(1.000, 10, by = 1))) +
        # coord_trans(y = "log10", ylim = c(0.001, 5)) +
        # scale_y_continuous(breaks = c(0.001, 0.01, 0.1, 1, 2, 3, 4))+
        #coord_cartesian(ylim = c(NA, 20)) +
        labs(col = "Dexa") + 
        #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
        #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
        # scale_color_manual(breaks = c("D+", "D-", "mild"),
        #                   values = c("orangered", "blue", "green3")) +
        # scale_fill_manual(breaks = c("D+", "D-", "mild"),
        #                   values = c("orangered", "blue", "green3")) +
        scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
        scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
        facet_wrap(~factor(Marker, levels = c("CD4 CD40L+ CD137+"#,
                                               #"CD4 CD40L+ TNF-a+","CD4 CD40L+ IL-21+"
                                              )),
                   ncol = 3) +
        geom_smooth(method = "rlm", fullrange = F, 
                    aes(group = Dexa, fill = Dexa), 
                    alpha = 0.1, 
                    se = T, size = 0.4) +
        # geom_line(position = position_dodge(0.4), aes(group = factor(PatID_Studie1)), 
        #         alpha = 0.3, size = 0.2)+
        #guides(fill = FALSE) +
        xlab("Time point") +
        ggtitle(paste0(i, " progression in CD4 cells")) +
        theme_classic() +
      theme(axis.line = element_line(colour = "black"),
      panel.grid.major = element_line(),
      panel.grid.minor = element_line(colour = "gray99"),
      panel.border = element_blank(),
      panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)))
  ggsave(paste0("MSimages/", i, " progressionCD4.png"), width = 10, height = 10, units = "cm", dpi = 300)
  ggsave(paste0("MSimages/", i, " progressionCD4.pdf"), width = 10, height = 10, units = "cm", dpi = 300)
  #ggsave(paste0("MSimages/SRTtimeseries.png"), width = 20, height = 10, units = "cm", dpi = 300)
}


```

<!-- ```{r} -->
<!-- ## SRT TagPSO -->
<!-- ## CD4 -->

<!-- # for(i in SRT) -->
<!-- # { -->
<!-- #   print( -->
<!-- #     facs %>% -->
<!-- #       filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% -->
<!-- #       filter(Marker == "CD4 CD40L+ CD137+" | Marker ==  "CD4 CD40L+ TNF-a+" | Marker == "CD4 CD40L+ IL-21+") %>% -->
<!-- #       #left_join(SRT_2_tp, ., by = "PatID_Studie1") %>% -->
<!-- #     ggplot(aes(x = TagPSO, y = !!sym(i), -->
<!-- #                   colour = Dexa)) + -->
<!-- #       geom_point(position = position_dodge(0.4), alpha = 0.5) + -->
<!-- #       geom_smooth(method = "rlm", -->
<!-- #                   fullrange = F, -->
<!-- #                   geom = "ribbon", -->
<!-- #                   aes(group = Dexa, fill = Dexa), -->
<!-- #                   alpha = 0.1, -->
<!-- #                   se = T, size = 0.4) + -->
<!-- #       labs(col = "Dexa") + -->
<!-- #       scale_y_log10(minor = c(seq(0.001, 0.010, by = 0.001), -->
<!-- #                               seq(0.010, 0.100, by = 0.01), -->
<!-- #                               seq(0.100, 1.000, by = 0.1), -->
<!-- #                               seq(1.000, 10, by = 1))) + -->
<!-- #       # coord_trans(y = "log10", ylim = c(0.001, 5)) + -->
<!-- #       # scale_y_continuous(breaks = c(0.001, 0.01, 0.1, 1, 2, 3, 4))+ -->
<!-- #       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!-- #                           values = c("orangered", "blue", "green3")) + -->
<!-- #       scale_fill_manual(breaks = c("D+", "D-", "mild"), -->
<!-- #                           values = c("orangered", "blue", "green3")) + -->
<!-- #       facet_wrap(~factor(Marker, levels = c("CD4 CD40L+ CD137+", -->
<!-- #                                              "CD4 CD40L+ TNF-a+","CD4 CD40L+ IL-21+")), -->
<!-- #                  ncol = 3) + -->
<!-- #       # geom_line(position = position_dodge(0.4), aes(group = factor(PatID_Studie1)), -->
<!-- #       #           alpha = 0.5, size = 0.2)+ -->
<!-- #       xlab("Days PSO") + -->
<!-- #       #guides(fill = FALSE) + -->
<!-- #       ggtitle(paste0(i, " progression in CD4 cells")) + -->
<!-- #       #theme(axis.text.x = element_text(angle = 90)) + -->
<!-- #       theme_classic() + -->
<!-- #       theme(axis.line = element_line(colour = "black"), -->
<!-- #       panel.grid.major = element_line(), -->
<!-- #       panel.grid.minor = element_line(colour = "gray99"), -->
<!-- #       panel.border = element_blank(), -->
<!-- #       panel.background = element_blank())+ -->
<!-- #       theme(axis.text = element_text(size = 11)) + -->
<!-- #       theme(strip.background =element_rect(fill="whitesmoke", colour = "white"), -->
<!-- #             strip.text.x = element_text(size = 11))) -->
<!-- #   ggsave(paste0("MSimages/", i, " progressionCD4_DaysPSO.png"), width = 20, height = 10, units = "cm", dpi= 300) -->
<!-- # } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(i in SRT) -->
<!-- {   -->
<!--   print( -->
<!--     facs %>% -->
<!--       filter(Marker1 == "CD4") %>% -->
<!--       filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%  -->
<!--       filter(Marker2 == "CD40L+ CD137+" | Marker2 ==  "CD40L+ TNF-a+" | Marker2 == "CD40L+ IL-21+") %>%  -->
<!--       #left_join(SRT_2_tp, ., by = "PatID_Studie1") %>%  -->
<!--     ggplot(., aes(x = factor(Timepoint,  -->
<!--                              levels = c('W2', 'M1', 'M3',  -->
<!--                                         'M6+')),  -->
<!--                   y = !!sym(i),  -->
<!--                   colour = Dexa, -->
<!--                   group = factor(PatID_Studie1))) + -->
<!--       geom_point(position = position_dodge(0.4), alpha = 0.5) + -->
<!--       #geom_jitter(width = 0.1, alpha = 0.2) + -->
<!--       coord_trans(y = "log10", ylim = c(0.001, 5)) + -->
<!--       scale_y_continuous(breaks = c(0.001, 0.01, 0.1, 1, 2, 3, 4, 5))+ -->
<!--       labs(col = "Dexa") +  -->
<!--       #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,  -->
<!--       #             colour = "gray47", alpha = 0.5, outlier.shape = NA) + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values = c("orangered", "blue", "green3")) + -->
<!--       scale_fill_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values = c("orangered", "blue", "green3")) + -->
<!--       geom_smooth(method = "lm", aes(group = Dexa, fill = Dexa), alpha = 0.1,  -->
<!--                   se = F, size = 0.4) +  -->
<!--       geom_line(position = position_dodge(0.4), aes(group = factor(PatID_Studie1)),  -->
<!--                 alpha = 0.5, size = 0.2)+ -->
<!--       facet_grid(factor(Marker2, levels = c("CD40L+ CD137+", -->
<!--                                              "CD40L+ TNF-a+","CD40L+ IL-21+")) ~ factor(Dexa, levels = c('D+', 'D-', 'mild')), -->
<!--                  #ncol = 3, -->
<!--                  # labeller = labeller(Marker2 = as_labeller(marker.labels), -->
<!--                  #                     Dexa = as_labeller(dexa.labels)), -->
<!--                  scales = "free") + -->
<!--       xlab("Time point") + -->
<!--       ggtitle(paste0(i, " progression in CD4 cells")) + -->
<!--       theme_bw() + -->
<!--       theme(axis.text = element_text(size = 6.5)) + -->
<!--       theme(strip.background =element_rect(fill="aliceblue"))) -->
<!--   ggsave(paste0("MSimages/", i, " progressionCD4_byDexa.png"), width = 20, height = 15, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(i in SRT) -->
<!-- {   -->
<!--   print( -->
<!--     facs %>% -->
<!--       filter(Marker1 == "CD4") %>% -->
<!--       filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%  -->
<!--       filter(Marker2 == "CD40L+ CD137+" | Marker2 ==  "CD40L+ TNF-a+" | Marker2 == "CD40L+ IL-21+") %>%  -->
<!--       #left_join(SRT_2_tp, ., by = "PatID_Studie1") %>%  -->
<!--     ggplot(., aes(x = TagPSO,y = !!sym(i),  -->
<!--                   colour = Dexa, -->
<!--                   group = factor(PatID_Studie1))) + -->
<!--       theme(axis.text.x = element_text(angle = 90)) + -->
<!--       geom_point(position = position_dodge(0.4), alpha = 0.5) + -->
<!--       #geom_jitter(width = 0.1, alpha = 0.2) + -->
<!--       coord_trans(y = "log10", ylim = c(0.001, 5)) + -->
<!--       scale_y_continuous(breaks = c(0.001, 0.01, 0.1, 1, 2, 3, 4, 5))+ -->
<!--       labs(col = "Dexa") +  -->
<!--       #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,  -->
<!--       #             colour = "gray47", alpha = 0.5, outlier.shape = NA) + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values = c("orangered", "blue", "green3")) + -->
<!--       scale_fill_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values = c("orangered", "blue", "green3")) + -->
<!--       geom_smooth(method = "lm", aes(group = Dexa, fill = Dexa), alpha = 0.1,  -->
<!--                   se = F, size = 0.5) + -->
<!--        geom_line(position = position_dodge(0.4), aes(group = factor(PatID_Studie1)),  -->
<!--                 alpha = 0.5, size = 0.2)+ -->
<!--       facet_grid(factor(Marker2, levels = c("CD40L+ CD137+", -->
<!--                                              "CD40L+ TNF-a+","CD40L+ IL-21+")) ~ factor(Dexa, levels = c('D+', 'D-', 'mild')), -->
<!--                  #ncol = 3, -->
<!--                  # labeller = labeller(Marker2 = as_labeller(marker.labels), -->
<!--                  #                     Dexa = as_labeller(dexa.labels)), -->
<!--                  scales = "free") + -->
<!--       xlab("Days PSO") + -->
<!--       ggtitle(paste0(i, " progression in CD4 cells")) + -->
<!--       theme_bw() + -->
<!--       theme(axis.text = element_text(size = 6.5)) + -->
<!--       theme(strip.background =element_rect(fill="aliceblue"))) -->
<!--   ggsave(paste0("MSimages/", i, " progressionCD4_DaysPSO_byDexa.png"), width = 20, height = 15, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

### PNT time series

(No patient had more than one measurement in one time point)
```{r PNT time series}
## PNT
# B.1.1.7
i = "B.1.1.7"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
      filter(exclude_B.1.1.7 == 0) %>% 
      #mutate(exclude_col := "{exclude}_{{ i }}")
      #filter("{exclude}_{{ i }}" == 0) %>% 
      #left_join(PNT_2_tp, ., by = "PatID_Studie1") %>% 
      ggplot(., aes(x = factor(Timepoint, 
                             levels = c('W2', 'M1', 'M3', 
                                        'M6')), 
                  y = !!sym(i), 
                  colour = Dexa)) +
      geom_jitter(width = 0.1, alpha = 0.5) +
      labs(col = "Dexa") +
      scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000),
                        seq(10000,100000,by=10000))) +
      #coord_trans(y = "log10", ylim = c(5, 37000)) +
      #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
      #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
      #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
      scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
      #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
      #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
      #                                        "CD137+ IFN-g+","CD137+ IL-21+",
      #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
      #            scales = "free") +
      # facet_grid(. ~ factor(Dexa, levels = c('1', '0', '3')),
                 #ncol = 3,
                 #labeller = labeller(Dexa = as_labeller(dexa.labels))) +
      geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                  se = T, size = 0.4) +
      xlab("Time point") +
      ggtitle(paste0(i, " neutralisation progression")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)))
ggsave(paste0("MSimages/", i, " neutralisation progression.png"), width = 20, height = 15, units = "cm", dpi = 300)
ggsave(paste0("MSimages/", i, " neutralisation progression.pdf"), width = 20, height = 15, units = "cm", dpi = 300)

B117_age_sex_tp_dexa_minus <- lmerTest::lmer(data = B117, log10(B.1.1.7) ~ Timepoint + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

B117_age_sex_tp_dexa_mild <- lmerTest::lmer(data = B117, log10(B.1.1.7) ~ Timepoint + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(B117_age_sex_tp_dexa_minus, B117_age_sex_tp_dexa_mild)
```

```{r}
i = "B.1.617.2"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
      filter(exclude_B.1.617.2 == 0) %>% 
      #mutate(exclude_col := "{exclude}_{{ i }}")
      #filter("{exclude}_{{ i }}" == 0) %>% 
      #left_join(PNT_2_tp, ., by = "PatID_Studie1") %>% 
      ggplot(., aes(x = factor(Timepoint, 
                             levels = c('W2','M1', 'M3', 
                                        'M6')), 
                  y = !!sym(i), 
                  colour = Dexa)) +
      geom_jitter(width = 0.1, alpha = 0.5) +
      labs(col = "Dexa") +
      scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000),
                        seq(10000,100000,by=10000))) +
      #coord_trans(y = "log10", ylim = c(5, 37000)) +
      #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
      #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
      #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
      #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
      #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
      #                                        "CD137+ IFN-g+","CD137+ IL-21+",
      #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
      #            scales = "free") +
      # facet_grid(. ~ factor(Dexa, levels = c('1', '0', '3')),
                 #ncol = 3,
                 #labeller = labeller(Dexa = as_labeller(dexa.labels))) +
      geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                  se = T, size = 0.4) +
      xlab("Time point") +
      ggtitle(paste0(i, " neutralisation progression")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)))
ggsave(paste0("MSimages/", i, " neutralisation progression.png"), width = 20, height = 15, units = "cm", dpi = 300)
ggsave(paste0("MSimages/", i, " neutralisation progression.pdf"), width = 20, height = 15, units = "cm", dpi = 300)


B16172_age_sex_tp_dexa_minus <- lmerTest::lmer(data = B16172, log10(B.1.617.2) ~ Timepoint + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

B16172_age_sex_tp_dexa_mild <- lmerTest::lmer(data = B16172, log10(B.1.617.2) ~ Timepoint + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(B16172_age_sex_tp_dexa_minus, B16172_age_sex_tp_dexa_mild)
```


```{r}
i = "B.1.1.7"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
      filter(exclude_B.1.1.7 == 0) %>% 
      #left_join(PNT_2_tp, ., by = "PatID_Studie1") %>% 
      ggplot(., aes(x = TagPSO, y = !!sym(i), 
                  colour = Dexa)) +
      geom_jitter(width = 0.1, alpha = 0.5) +
      labs(col = "Dexa") +
      scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000),
                        seq(10000,100000,by=10000))) +
      #coord_trans(y = "log10", ylim = c(5, 37000)) +
      #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
      #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
      #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
      #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
      #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
      #                                        "CD137+ IFN-g+","CD137+ IL-21+",
      #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
      #            scales = "free") +
      # facet_wrap(. ~ factor(Dexa, levels = c('D+', 'D-', 'mild')),
      #            #ncol = 3,
      #            scales = "free",
      #            labeller = labeller(Dexa = as_labeller(dexa.labels))) +
      geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                  se = T, size = 0.4, fullrange = F) +
      xlab("Days PSO") +
      ggtitle(paste0(i, " neutralisation progression")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)))
ggsave(paste0("MSimages/", i, " neutralisation progression DaysPSO.png"), width = 20, height = 15, units = "cm", dpi = 300)

B117_age_sex_tPSO_dexa_minus <- lmerTest::lmer(data = B117, log10(B.1.1.7) ~ TagPSO + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

B117_age_sex_tPSO_dexa_mild <- lmerTest::lmer(data = B117, log10(B.1.1.7) ~ TagPSO + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(B117_age_sex_tp_dexa_minus, B117_age_sex_tp_dexa_mild)
# https://stats.stackexchange.com/questions/435644/is-there-a-method-to-look-for-significant-difference-between-two-linear-regressi
# https://stackoverflow.com/questions/33192618/looking-for-differences-btw-linear-regression-lines-in-r
```
  
```{r}  
i = "B.1.617.2"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
      filter(exclude_B.1.617.2 == 0) %>% 
      #left_join(PNT_2_tp, ., by = "PatID_Studie1") %>% 
      ggplot(., aes(x = TagPSO, y = !!sym(i), 
                  colour = Dexa)) +
      geom_jitter(width = 0.1, alpha = 0.5) +
      labs(col = "Dexa") +
      scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000),
                        seq(10000,100000,by=10000))) +
      #coord_trans(y = "log10", ylim = c(5, 37000)) +
      #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
      #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
      #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
      #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
      #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
      #                                        "CD137+ IFN-g+","CD137+ IL-21+",
      #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
      #            scales = "free") +
      # facet_wrap(. ~ factor(Dexa, levels = c('D+', 'D-', 'mild')),
      #            #ncol = 3,
      #            scales = "free",
      #            labeller = labeller(Dexa = as_labeller(dexa.labels))) +
      geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                  se = T, size = 0.4, fullrange = F) +
      xlab("Days PSO") +
      ggtitle(paste0(i, " neutralisation progression")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)))
ggsave(paste0("MSimages/", i, " neutralisation progression DaysPSO.png"), width = 20, height = 15, units = "cm", dpi = 300)

B16172_age_sex_tPSO_dexa_minus <- lmerTest::lmer(data = B16172, log10(B.1.617.2) ~ TagPSO + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

B16172_age_sex_tPSO_dexa_mild <- lmerTest::lmer(data = B16172, log10(B.1.617.2) ~ TagPSO + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(B16172_age_sex_tp_dexa_minus, B16172_age_sex_tp_dexa_mild)
```


```{r}

## PNT separated by dexa and joined by patients

i = "B.1.1.7"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
      filter(exclude_B.1.1.7 == 0) %>%
      ggplot(aes(x = factor(Timepoint,
                            levels = c('W2', 'M1', 'M3',
                                          'M6')),
                 y = !!sym(i),
                 colour = Dexa,
                 group = factor(PatID_Studie1))) +
        labs(col = "Dexa") +
        theme(axis.text.x = element_text(angle = 90)) +
        #geom_jitter(width = 0.1, alpha = 0.5) +
        geom_point(position = position_dodge(0.1), alpha = 0.5) +
        scale_y_log10(minor=c(5, 50, 500, 5000, 50000)) +
      # minor=c(1:10,
      #                   seq( 10, 100,by=10 ),
      #                   seq(100,1000,by=100),
      #                   seq(1000,10000,by=1000),
      #                   seq(10000,100000,by=10000))
        #coord_trans(y = "log10") +
        #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+

        #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,
        #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
        scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
        #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
        #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
        #                                        "CD137+ IFN-g+","CD137+ IL-21+",
        #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
        #            scales = "free") +
        # facet_wrap(. ~ factor(Timepoint, levels = c('D+', 'D-', 'mild')),
        #          ncol = 3,
        #          scales = "free")+
                 #labeller = labeller(Dexa = as_labeller(dexa.labels))) +
        geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1,
                  se = T, size = 0.7) +
        geom_line(position = position_dodge(0.1),
                  aes(group = factor(PatID_Studie1)), alpha = 0.5,
                  size = 0.2) +
        xlab("Time point") +
        ggtitle(paste0(i, " neutralisation progression")) +
        theme_bw() +
        #theme(legend.position = "none") +
        theme(strip.background =element_rect(fill="aliceblue")))
ggsave(paste0("MSimages/", i, " neutralisation progression DaysPSO byPatient.png"), width = 20, height = 10, units = "cm", dpi = 300)

i = "B.1.617.2"
print(
    PNT %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
      filter(exclude_B.1.617.2 == 0) %>%
      ggplot(aes(x = factor(Timepoint,
                            levels = c('W2', 'M1', 'M3',
                                          'M6')),
                 y = !!sym(i),
                 colour = Dexa,
                 group = factor(PatID_Studie1))) +
        labs(col = "Dexa") +
        theme(axis.text.x = element_text(angle = 90)) +
        #geom_jitter(width = 0.1, alpha = 0.5) +
        geom_point(position = position_dodge(0.1), alpha = 0.5) +
        scale_y_log10(minor=c(5, 50, 500, 5000, 50000)) +
      # minor=c(1:10,
      #                   seq( 10, 100,by=10 ),
      #                   seq(100,1000,by=100),
      #                   seq(1000,10000,by=1000),
      #                   seq(10000,100000,by=10000))
        #coord_trans(y = "log10") +
        #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+

        #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,
        #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
        scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
        #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
        #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
        #                                        "CD137+ IFN-g+","CD137+ IL-21+",
        #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
        #            scales = "free") +
        # facet_wrap(. ~ factor(Timepoint, levels = c('D+', 'D-', 'mild')),
        #          ncol = 3,
        #          scales = "free")+
                 #labeller = labeller(Dexa = as_labeller(dexa.labels))) +
        geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1,
                  se = T, size = 0.7) +
        geom_line(position = position_dodge(0.1),
                  aes(group = factor(PatID_Studie1)), alpha = 0.5,
                  size = 0.2) +
        xlab("Time point") +
        ggtitle(paste0(i, " neutralisation progression")) +
        theme_bw() +
        #theme(legend.position = "none") +
        theme(strip.background =element_rect(fill="aliceblue")))
ggsave(paste0("MSimages/", i, " neutralisation progression DaysPSO byPatient.png"), width = 20, height = 10, units = "cm", dpi = 300)

```

### IgG time series
(No patient had more than one measurement in one time point)

```{r IgG time series}
sero %>%
  filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
  ggplot(., aes(x = factor(Timepoint, 
                           levels = c('W2', 'M1', 'M3', 
                                      'M6')), 
                y = IgG, 
                colour = factor(Dexa))) +
    theme(axis.text.x = element_text(angle = 90)) +
    geom_jitter(width = 0.1, alpha = 0.5) +
    labs(col = "Dexa") +
    scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000))) +
  # coord_trans(y = "log10", ylim = c(5, 12000)) +
  # scale_y_continuous(breaks = c(10, 100, 1000,10000))+
      
    #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
    #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
    scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
    scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "lightslateblue", "black")) +
    # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
    #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
    #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
    #                                        "CD137+ IFN-g+","CD137+ IL-21+",
    #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
    #            scales = "free") +
    geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                se = T, size = 0.4) +
    xlab("Time point") +
    ggtitle("IgG progression") +
    theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11))
ggsave(paste0("MSimages/IgG progression.png"), width = 20, height = 15, units = "cm", dpi = 300)
ggsave(paste0("MSimages/IgG progression.pdf"), width = 20, height = 15, units = "cm", dpi = 300)


IgG_age_sex_tp_dexa_minus <- lmerTest::lmer(data = sero_m, log10(IgG) ~ Timepoint + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

IgG_age_sex_tp_dexa_mild <- lmerTest::lmer(data = sero_m, log10(IgG) ~ Timepoint + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(IgG_age_sex_tp_dexa_minus, IgG_age_sex_tp_dexa_mild)
```

```{r}
sero %>%
  filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
  ggplot(aes(x = as.numeric(TagPSO), 
                y = IgG, 
                colour = Dexa)) +
    theme(axis.text.x = element_text(angle = 90)) +
    geom_jitter(width = 0.1, alpha = 0.5) +
    labs(col = "Dexa") +
    scale_y_log10(minor=c(1:10,
                        seq( 10, 100,by=10 ),
                        seq(100,1000,by=100),
                        seq(1000,10000,by=1000))) +
    #coord_trans(y = "log10", ylim = c(5, 12000)) +
    #scale_y_continuous(breaks = c(10, 100, 1000,10000))+
    #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5, 
    #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
    scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
    scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
    # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
    #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
    #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
    #                                        "CD137+ IFN-g+","CD137+ IL-21+",
    #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
    #            scales = "free") +
    geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1, 
                se = T, size = 0.4) +
    xlab("Days PSO") +
    ggtitle("IgG progression") +
    theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11))
ggsave(paste0("MSimages/IgG progression DaysPSO.png"), width = 20, height = 15, units = "cm", dpi = 300)

IgG_age_sex_tPSO_dexa_minus <- lmerTest::lmer(data = sero_m, log10(IgG) ~ TagPSO + relevel(as.factor(Dexa), ref = "D-") + Age + Sex + (1|PatID_Studie1))

#tab_model(B117_age_sex_tp_dexa_minus)

IgG_age_sex_tPSO_dexa_mild <- lmerTest::lmer(data = sero_m, log10(IgG) ~ TagPSO + relevel(as.factor(Dexa), ref = "mild") + Age + Sex + (1|PatID_Studie1))

tab_model(IgG_age_sex_tPSO_dexa_minus, IgG_age_sex_tPSO_dexa_mild)
```

```{r}
# sero by patient
sero %>%
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
      ggplot(aes(x = factor(Timepoint,
                            levels = c('W2', 'M1', 'M3',
                                          'M6')),
                 y = IgG,
                 colour = Dexa,
                 group = factor(PatID_Studie1))) +
        labs(col = "Dexa") +
        theme(axis.text.x = element_text(angle = 90)) +
        #geom_jitter(width = 0.1, alpha = 0.5) +
        geom_point(position = position_dodge(0.1), alpha = 0.5) +
        scale_y_log10(minor=c(5, 50, 500, 5000, 50000)) +
      # minor=c(1:10,
      #                   seq( 10, 100,by=10 ),
      #                   seq(100,1000,by=100),
      #                   seq(1000,10000,by=1000),
      #                   seq(10000,100000,by=10000))
        #coord_trans(y = "log10") +
        #scale_y_continuous(breaks = c(10, 100, 1000,10000, 20000, 30000, 40000))+

        #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,
        #             colour = "gray47", alpha = 0.5, outlier.shape = NA) +
        scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        scale_fill_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
        # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+",
        #                                        "CD40L+ IFN-g+","CD40L+ IL-21+",
        #                                        "CD40L+ TNF-a+","CD137+ GrzB+",
        #                                        "CD137+ IFN-g+","CD137+ IL-21+",
        #                                        "CD137+ TNF-a+","CD137+ GrzB-")),
        #            scales = "free") +
        # facet_wrap(. ~ factor(Timepoint, levels = c('D+', 'D-', 'mild')),
        #          ncol = 3,
        #          scales = "free")+
                 #labeller = labeller(Dexa = as_labeller(dexa.labels))) +
        geom_smooth(method = "rlm", aes(group = Dexa, fill = Dexa), alpha = 0.1,
                  se = T, size = 0.7) +
        geom_line(position = position_dodge(0.1),
                  aes(group = factor(PatID_Studie1)), alpha = 0.5,
                  size = 0.2) +
        xlab("Time point") +
        ggtitle("IgG progression") +
        theme_bw() +
        #theme(legend.position = "none") +
        theme(strip.background =element_rect(fill="aliceblue"))
ggsave(paste0("MSimages/IgG progression DaysPSO ByPatient.png"), width = 20, height = 15, units = "cm", dpi = 300)

```

# Correlations
(Plots are made in log scale but Spearman's rank correlations were performed on raw data)

### IgG vs PNT correlation
```{r Association of IgG and PNT at the respective timepoints}

Sero_PNT <- sero %>% 
  dplyr::select(PatID_Studie1, Dexa, Timepoint, IgG) %>% 
  full_join(., PNT, by = c("PatID_Studie1", "Dexa", "Timepoint"))

# IgG vs B.1.1.7

Sero_PNT %>% 
  filter(!Timepoint == "PI") %>% 
  #filter((Timepoint == "M3" | Timepoint == "M6")) %>% 
  ggplot(aes(x = B.1.1.7, y = IgG)) +
  geom_point(alpha = 0.7, aes(colour = factor(Dexa))) +
  labs(col = "Dexa") +
  coord_trans(y = "log10") +
  scale_y_continuous(breaks = c(10, 100, 1000))+
  scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
  facet_wrap(~factor(Timepoint, levels = c('W2', 'M1', 'M3','M6')), 
             scales = "free_x",
             drop = T) +
  theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
  ggtitle("IgG vs PNT measures at the same time point") +
  xlab("B.1.1.7") + ylab("IgG BAU/ml") +
  stat_cor(method = "spearman", size = 3) 
  #geom_smooth(method = "lm", se = T)
ggsave(paste0("MSimages/IgG vs B.1.1.7.png"), width = 20, height = 10, units = "cm", dpi = 300)
```

```{r}
Sero_PNT %>% 
  filter(!Timepoint == "PI")  %>% 
  #filter((Timepoint == "M3" | Timepoint == "M6+"))%>% 
  ggplot(aes(x = B.1.1.7, y = IgG, colour = factor(Dexa))) +
  geom_point(alpha = 0.7) +
  coord_trans(y = "log10") +
  scale_y_continuous(breaks = c(10, 100, 1000))+
  labs(col = "Dexa") +
  scale_color_manual(breaks = c("D+", "D-", "mild"),
                      values = c("orangered", "blue", "green3")) +
  facet_wrap(~factor(Timepoint, levels = c('W2', 'M1', 'M3', 'M6')), 
             scales = "free_x",
             ncol = 2) +
  theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
  ggtitle("IgG vs PNT measures at the same time point") +
  xlab("B.1.1.7") + ylab("IgG BAU/ml") +
  stat_cor(method = "spearman", size = 3) #+
  #geom_smooth(method = "lm", se = T)
ggsave(paste0("MSimages/IgG vs B.1.1.7 slopeByDexa.png"), width = 20, height = 10, units = "cm")
```

```{r}
# IgG vs B.1.617.2

Sero_PNT %>% 
  filter(!Timepoint == "PI")  %>% 
  #filter((Timepoint == "M3" | Timepoint == "M6+"))%>% 
  ggplot(aes(x = B.1.617.2, y = IgG)) +
  geom_point(alpha = 0.7, aes(colour = factor(Dexa))) +
  coord_trans(y = "log10") +
  scale_y_continuous(breaks = c(10, 100, 1000))+
  labs(col = "Dexa") +
  scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
  facet_wrap(~factor(Timepoint, levels = c('W2', 'M1', 'M3', 'M6')), 
             scales = "free_x",
             ncol = 2) +
  theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
  ggtitle("IgG vs PNT measures at the same time point") +
  xlab("B.1.617.2") + ylab("IgG BAU/ml") +
  stat_cor(method = "spearman", size = 3) 
  #geom_smooth(method = "lm", se = T)
ggsave(paste0("MSimages/IgG vs B.1.617.2.png"), width = 20, height = 10, units = "cm")
```

```{r}
Sero_PNT %>% 
  filter(!Timepoint == "PI")  %>% 
  #filter((Timepoint == "M3" | Timepoint == "M6+"))%>% 
  ggplot(aes(x = B.1.617.2, y = IgG, colour = factor(Dexa))) +
  geom_point(alpha = 0.7) +
  coord_trans(y = "log10", ylim = c(5, 12000)) +
  scale_y_continuous(breaks = c(10, 100, 1000))+
  labs(col = "Dexa") +
  scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
  facet_wrap(~factor(Timepoint, levels = c('W2', 'M1', 'M3', 'M6')), 
             scales = "free_x",
             ncol = 2) +
 theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
  ggtitle("IgG vs PNT measures at the same time point") +
  xlab("B.1.617.2") + ylab("IgG BAU/ml") +
  stat_cor(method = "spearman", size = 3) #+
  #geom_smooth(method = "lm", se = T)
ggsave(paste0("MSimages/IgG vs B.1.617.2 slopeByDexa.png"), width = 20, height = 10, units = "cm")

```

### IgG vs SRT correlation
```{r Associatin of IgG and SRT at timepoints}
Sero_SRT <- sero %>%
  dplyr::select(PatID_Studie1, Dexa, Timepoint, IgG) %>% 
  full_join(facs, .)

print(
    Sero_SRT %>% 
      filter(Marker == "CD4 CD40L+ CD137+" | Marker ==  "CD4 CD40L+ TNF-a+" | Marker == "CD4 CD40L+ IL-21+") %>% 
      #filter(Marker1=="CD4") %>% 
      filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
      ggplot(aes(x = IgG, y = S1)) +
      geom_point(alpha = 0.7, aes(color = factor(Dexa))) +
      coord_trans(y = "log10") + #, ylim = c(0,10)
      scale_y_continuous(breaks = seq(0,1.2,0.2))+
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      facet_wrap(~Marker, scales ="free") +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ GrzB+",
      #                                        "CD137+ GrzB+")),
      #            scales = "free") +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
      ggtitle("IgG vs SRT (CD4)") +
      xlab("IgG BAU/ml") +
      stat_cor(method = "spearman", size = 3.5)) 
      #geom_smooth(method = "lm", se = T))
  ggsave(paste0("MSimages/IgG vs S1.png"), width = 20, height = 10, units = "cm", dpi = 300)
```

### PNT vs SRT correlation
```{r Association of PNT and SRT at timepoints, fig.height = 20}
PNT_SRT <- PNT %>%
  dplyr::select(PatID_Studie1, Dexa, Timepoint, B.1.1.7, B.1.617.2, exclude_B.1.1.7, exclude_B.1.617.2) %>% 
  full_join(facs, .)

# B.1.1.7 vs S1
print(PNT_SRT %>%
    filter(grepl(Marker, pattern = "CD4")) %>% 
    filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
    #filter(Marker2 == "CD137+ IL-21+") %>% 
    filter(exclude_B.1.1.7 == 0) %>% 
    as.data.frame() %>%
    ggplot(., aes(x = B.1.1.7, y = S1)) +
      geom_point(alpha = 0.7, aes(color = factor(Dexa))) +
      coord_trans(y = "log10") + #, ylim = c(0,5)
      scale_y_continuous(breaks = seq(0,3,0.05)) +
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) + 
      facet_wrap(~Marker, scales = "free", ncol = 2) +
      # facet_wrap(~factor(Marker2, levels = c("CD40L+ IFN-g+",
      #                                        "CD40L+ IL-21+",
      #                                        "CD137+ IL-21+")),
      #            scales = "free") +
      ggtitle(paste("B.1.1.7 vs S1")) +
      stat_cor(method = "spearman", size = 4))
ggsave("MSimages/B.1.1.7 vs S1.png", width = 20, height = 10, units = "cm", dpi = 300)
```

```{r, fig.height=20}
# B.1.1.7 vs S2
print(PNT_SRT %>%
    filter(grepl(Marker, pattern = "CD4")) %>%
    filter(!(Timepoint == "PI" | Timepoint == "VC")) %>% 
    #filter(Marker2 == "CD137+ IL-21+") %>% 
    filter(exclude_B.1.1.7 == 0) %>% 
    as.data.frame() %>%
    ggplot(., aes(x = B.1.1.7, y = S2)) +
      geom_point(alpha = 0.7, aes(color = factor(Dexa))) +
      coord_trans(y = "log10") +
      scale_y_continuous(breaks = seq(0,0.3, 0.05)) +
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
      facet_wrap(~Marker, scales = "free", ncol = 2) +
      # facet_wrap(~factor(Marker2, levels = c("CD137+ IFN-g+",
      #                                        "CD40L+ IL-21+",
      #                                        "CD137+ TNF-a")),
      #            scales = "free") +
      ggtitle(paste("B.1.1.7 vs S2")) +
      stat_cor(method = "spearman", size = 4))
ggsave("MSimages/B.1.1.7 vs S2.png", width = 20, height = 10, units = "cm", dpi = 300)
```

```{r}
# B.1.617.2 vs S1
print(PNT_SRT %>%
    #filter(Marker1 == "CD4") %>%
    filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
    filter(Marker == "CD4 CD40L+ CD137+" | Marker == "CD4 CD40L+ IFN-g+") %>%
    filter(exclude_B.1.617.2 == 0) %>%
    as.data.frame() %>%
    ggplot(., aes(x = B.1.617.2, y = S1)) +
      geom_point(alpha = 0.7, aes(color = factor(Dexa))) +

      coord_trans(y = "log10") +
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      theme_bw() +
      facet_wrap(~Marker, scales = "free", ncol = 2) +
      # facet_wrap(~factor(Marker2),#, levels = c("CD40L+ CD137+",
      #                                        #"CD40L+ IL-21+",
      #                                        #"CD40L+ TNF-a+")),
      #            scales = "free") +
      theme(strip.background =element_rect(fill="aliceblue")) +
      ggtitle(paste("B.1.617.2 vs S1")) +
      stat_cor(method = "spearman", size = 4))
ggsave(paste0("MSimages/B.1.617.2 vs S1.png"), width = 20, height = 10, units = "cm", dpi = 300)
```

```{r}
# B.1.617.2 vs S2
print(PNT_SRT %>%
    #filter(Marker1 == "CD4") %>%
    filter(!(Timepoint == "PI" | Timepoint == "VC")) %>%
    filter(Marker == "CD4 CD40L+ CD137+" | Marker == "CD8 CD137+ GrzB-") %>%
    filter(exclude_B.1.617.2 == 0) %>%
    as.data.frame() %>%
    ggplot(., aes(x = B.1.617.2, y = S2)) +
      geom_point(alpha = 0.7, aes(color = factor(Dexa))) +

      coord_trans(y = "log10") +
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                          values = c("orangered", "blue", "green3")) +
      theme_bw() +
      facet_wrap(~Marker, scales = "free", ncol = 2) +
      # facet_wrap(~factor(Marker2),#, levels = c("CD40L+ CD137+",
      #                                        #"CD40L+ IL-21+",
      #                                        #"CD40L+ TNF-a+")),
      #            scales = "free") +
      theme(strip.background =element_rect(fill="aliceblue")) +
      ggtitle(paste("B.1.617.2 vs S2")) +
      stat_cor(method = "spearman", size = 4))
ggsave(paste0("MSimages/B.1.617.2 vs S2.png"), width = 10, height = 10, units = "cm", dpi = 300)

```

## Early vs late timepoints
### SRT at W2 vs IgG at M6
```{r}
# S1 @ W2 vs IgG @ M6+: CD4 CD40L+ GrzB+
SRT_W2_sero_M5_pat <- sero %>% 
  left_join(., facs, by = "PatID_Studie1") %>% 
  group_by(PatID_Studie1) %>% 
  rename(Sero_time = Timepoint.x, SRT_time = Timepoint.y) %>% 
  filter(Sero_time == "M6", SRT_time == "W2")

print(
    SRT_W2_sero_M5_pat %>% 
      filter(Marker == "CD4 CD40L+ GrzB+") %>% 
      ggplot(aes(x = IgG, y = S1)) + 
      geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) + 
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                        values=c("orangered", "blue", "green3")) +
      facet_wrap(~factor(Marker, 
                         levels = c("CD4 CD40L+ GrzB+")),
                 scales = "free_y", ncol = 2) +
      stat_cor(method = "spearman", size = 4) +
      #stat_smooth(method = "lm", se = "F") +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
      xlab("IgG @ M6") +
      ylab(paste0("S1 @ W2")) +
      ggtitle("CD4 early SRT vs later IgG")
  )
ggsave(paste("MSimages/IgG@M6 vs SRT@W2 CD40L+ GrzB+.png"), width = 20, height = 20, units = "cm", dpi = 300)

```

<!-- ```{r early SRT vs late IgG} -->

<!-- # SRT W2 vs IgG after month 5 -->
<!-- SRT_W2_sero_M5_pat <- sero %>%  -->
<!--   left_join(., facs, by = "PatID_Studie1") %>%  -->
<!--   group_by(PatID_Studie1) %>%  -->
<!--   rename(Sero_time = Timepoint.x, SRT_time = Timepoint.y) %>%  -->
<!--   filter(Sero_time == "M6+", SRT_time == "W2") -->

<!-- for(i in SRT) -->
<!-- {  print( -->
<!--     SRT_W2_sero_M5_pat %>%  -->
<!--       filter(Marker1 == "CD4") %>%  -->
<!--       ggplot(aes(x = IgG, y = !!sym(i))) +  -->
<!--       geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +  -->
<!--       labs(col = "Dexa") + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                         values=c("orangered", "blue", "green3")) + -->
<!--       facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--                                              "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--                                              "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--                                              "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--                                              "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--                  scales = "free_y", ncol = 2) + -->
<!--       stat_cor(method = "spearman", size = 4) + -->
<!--       #stat_smooth(method = "lm", se = "F") + -->
<!--       theme_bw() + -->
<!--       theme(strip.background =element_rect(fill="aliceblue")) + -->
<!--       xlab("IgG @ M6+") + -->
<!--       ylab(paste0(i, " @ W2")) + -->
<!--       ggtitle("CD4 early SRT vs later IgG") -->
<!--   ) -->
<!-- ggsave(paste("MSimages/IgG@M6+ vs CD4 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(i in SRT) -->
<!-- {  print( -->
<!--     SRT_W2_sero_M5_pat %>%  -->
<!--       filter(Marker1 == "CD8") %>%  -->
<!--       ggplot(aes(x = IgG, y = !!sym(i))) +  -->
<!--       geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +  -->
<!--       labs(col = "Dexa") + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                         values=c("orangered", "blue", "green3")) + -->
<!--       facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--                                              "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--                                              "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--                                              "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--                                              "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--                  scales = "free_y", ncol = 2) + -->
<!--       stat_cor(method = "spearman", size = 4) + -->
<!--       #stat_smooth(method = "lm", se = "F") + -->
<!--       theme_bw() + -->
<!--       theme(strip.background =element_rect(fill="aliceblue")) + -->
<!--       xlab("IgG @ M6+") + -->
<!--       ylab(paste0(i, " @ W2")) + -->
<!--       ggtitle("CD8 early SRT vs later IgG") -->
<!--   ) -->
<!-- ggsave(paste("MSimages/IgG@M6+ vs CD8 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

### SRT at W2 vs PNT at M6
```{r}

PNT_SRT_diff_timepoints <- PNT %>% 
  left_join(., facs, by = "PatID_Studie1") %>% 
  group_by(PatID_Studie1) %>% 
  rename(SRT_time = Timepoint.y) %>% 
  rename(PNT_time = Timepoint.x)

print(
    PNT_SRT_diff_timepoints %>% 
      filter(!(Dexa.x == "4")) %>% 
      #filter(PNT_time == "M6+", SRT_time == "W2") %>% 
      filter(Marker == "CD4 CD40L+ IFN-g+" | Marker == "CD4 CD40L+ IL-21+") %>% 
      filter(exclude_B.1.617.2 == 0) %>% 
      ggplot(aes(x = B.1.617.2, y = S1)) + 
      geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) + 
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                        values=c("orangered", "blue", "green3")) +
      facet_wrap(~Marker, 
                 #levels = c("CD137+ IFN-g+")),
                 scales = "free_y", ncol = 2) +
      stat_cor(method = "spearman", size = 4) +
      #stat_smooth(method = "lm", se = "F") +
      theme_classic() +
      theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_line(),
    panel.grid.minor = element_line(colour = "gray99"),
    panel.border = element_blank(),
    panel.background = element_blank())+
      theme(axis.text = element_text(size = 11)) +
      theme(strip.background =element_rect(fill="whitesmoke", colour = "white"),
            strip.text.x = element_text(size = 11)) +
      xlab("PNT (B.1.617.2) @ M6") +
      ylab(paste0("S1 @ W2")) +
      ggtitle("CD4 early SRT vs later PNT")
  )
ggsave(paste("MSimages/B.1.617.2@M6+ vs SRT@W2.png"), width = 20, height = 20, units = "cm", dpi = 300)

```

```{r early SRT vs late PNT, fig.height = 20}

PNT_SRT_diff_timepoints <- PNT %>%
  left_join(., facs, by = "PatID_Studie1") %>%
  group_by(PatID_Studie1) %>%
  rename(SRT_time = Timepoint.y) %>%
  rename(PNT_time = Timepoint.x)

for(i in SRT)
 { print(
    PNT_SRT_diff_timepoints %>%
      #filter(PNT_time == "M6+", SRT_time == "W2") %>%
      filter(grepl(Marker, pattern ="CD4")) %>%
      filter(Dexa.x != "4") %>% 
      filter(exclude_B.1.1.7 == 0) %>%
      ggplot(aes(x = B.1.1.7, y = !!sym(i))) +
      geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +
      labs(col = "Dexa") +
      scale_color_manual(breaks = c("D+", "D-", "mild"),
                        values=c("orangered", "blue", "green3")) +
      facet_wrap(~Marker, 
                         # levels = c("CD40L+ CD137+","CD40L+ GrzB+",
                         #                     "CD40L+ IFN-g+","CD40L+ IL-21+",
                         #                     "CD40L+ TNF-a+","CD137+ GrzB+",
                         #                     "CD137+ IFN-g+","CD137+ IL-21+",
                         #                     "CD137+ TNF-a+","CD137+ GrzB-")),
      scales = "free_y", ncol = 2) +
      stat_cor(method = "spearman", size = 4) +
      #stat_smooth(method = "lm", se = "F") +
      theme_bw() +
      theme(strip.background =element_rect(fill="aliceblue")) +
      xlab("PNT (B.1.1.7) @ M6") +
      ylab(paste0(i, " @ W2")) +
      ggtitle("CD4 early SRT vs later PNT")
  )
ggsave(paste("MSimages/B.1.1.7@M6+ vs CD4 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300)
}
```

<!-- ```{r} -->
<!-- for(i in SRT) -->
<!--  { print( -->
<!--     PNT_SRT_diff_timepoints %>%  -->
<!--       filter(PNT_time == "M6+", SRT_time == "W2") %>%  -->
<!--       filter(Marker1 == "CD8") %>%  -->
<!--       filter(exclude_B.1.1.7 == 0) %>%  -->
<!--       ggplot(aes(x = B.1.1.7, y = !!sym(i))) +  -->
<!--       geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +  -->
<!--       labs(col = "Dexa") + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                         values=c("orangered", "blue", "green3")) + -->
<!--       facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--                                              "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--                                              "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--                                              "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--                                              "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--                  scales = "free_y", ncol = 2) + -->
<!--       stat_cor(method = "spearman", size = 4) + -->
<!--       #stat_smooth(method = "lm", se = "F") + -->
<!--       theme_bw() + -->
<!--       theme(strip.background =element_rect(fill="aliceblue")) + -->
<!--       xlab("PNT (B.1.1.7) @ M6+") + -->
<!--       ylab(paste0(i, " @ W2")) + -->
<!--       ggtitle("CD8 early SRT vs later PNT") -->
<!--   ) -->
<!-- ggsave(paste("MSimages/B.1.1.7@M6+ vs CD8 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # B.1.617.2 -->
<!-- for(i in SRT) -->
<!--  { print( -->
<!--     PNT_SRT_diff_timepoints %>%  -->
<!--       filter(PNT_time == "M6+", SRT_time == "W2") %>%  -->
<!--       filter(Marker1 == "CD4") %>%  -->
<!--       filter(exclude_B.1.617.2 == 0) %>%  -->
<!--       ggplot(aes(x = B.1.617.2, y = !!sym(i))) +  -->
<!--       geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +  -->
<!--       labs(col = "Dexa") + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                         values=c("orangered", "blue", "green3")) + -->
<!--       facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--                                              "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--                                              "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--                                              "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--                                              "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--                  scales = "free_y", ncol = 2) + -->
<!--       stat_cor(method = "spearman", size = 4) + -->
<!--       #stat_smooth(method = "lm", se = "F") + -->
<!--       theme_bw() + -->
<!--       theme(strip.background =element_rect(fill="aliceblue")) + -->
<!--       xlab("PNT (B.1.617.2) @ M6+") + -->
<!--       ylab(paste0(i, " @ W2")) + -->
<!--       ggtitle("CD4 early SRT vs later PNT") -->
<!--   ) -->
<!-- ggsave(paste("MSimages/B.1.617.2@M6+ vs CD4 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(i in SRT) -->
<!--  { print( -->
<!--     PNT_SRT_diff_timepoints %>%  -->
<!--       filter(PNT_time == "M6+", SRT_time == "W2") %>%  -->
<!--       filter(Marker1 == "CD8") %>%  -->
<!--       filter(exclude_B.1.617.2 == 0) %>%  -->
<!--       ggplot(aes(x = B.1.617.2, y = !!sym(i))) +  -->
<!--       geom_point(alpha = 0.7, aes(colour = factor(Dexa.x))) +  -->
<!--       labs(col = "Dexa") + -->
<!--       scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                         values=c("orangered", "blue", "green3")) + -->
<!--       facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--                                              "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--                                              "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--                                              "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--                                              "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--                  scales = "free_y", ncol = 2) + -->
<!--       stat_cor(method = "spearman", size = 4) + -->
<!--       #stat_smooth(method = "lm", se = "F") + -->
<!--       theme_bw() + -->
<!--       theme(strip.background =element_rect(fill="aliceblue")) + -->
<!--       xlab("PNT (B.1.617.2) @ M6+") + -->
<!--       ylab(paste0(i, " @ W2")) + -->
<!--       ggtitle("CD8 early SRT vs later PNT") -->
<!--   ) -->
<!-- ggsave(paste("MSimages/B.1.617.2@M6+ vs CD8 ", i, "@W2.png"), width = 20, height = 20, units = "cm", dpi = 300) -->
<!-- } -->

<!-- ``` -->

<!-- ```{r Evolution of T-cells, PNT and IgG before and after dexa} -->

<!-- # serology -->

<!-- sero_valid_dates <- sero %>%  -->
<!--   filter(!is.na(`#Ende Dexa`)) -->

<!-- # make it wide -->
<!-- print( -->
<!--   sero_evol <- sero_valid_dates %>%  -->
<!--   # select(Studie1, -->
<!--   #        PatID_Studie1, -->
<!--   #        Dexa, -->
<!--   #        Timepoint, -->
<!--   #        Marker2, -->
<!--   #        S1) %>% -->
<!--   #group_by(PatID_Studie1) %>% -->
<!--   mutate(row = row_number()) %>% -->
<!--   mutate(Entnahmedatum = gsub(".", "-", Entnahmedatum, fixed = T)) %>%  -->
<!--   mutate(`#Ende Dexa` = gsub(".", "-", `#Ende Dexa`, fixed = T)) %>% -->
<!--   mutate(`#Start Dexa` = gsub(".", "-", `#Start Dexa`, fixed = T)) %>% -->
<!--   mutate(End_to_sample = as.numeric(as.Date(Entnahmedatum, format = "%d-%m-%Y") - as.Date(`#Ende Dexa`, format = "%d-%m-%Y"))) %>%  -->
<!--   mutate(Dexa_end_zero = rep(0,nrow(.))) %>%  -->
<!--   # tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, Dexa,  -->
<!--   #                                `#Start Dexa`, `#Ende Dexa`,  -->
<!--   #                                `#Days Start Dexa PSO`, `#Duration Dexa`), -->
<!--   #                    names_from = Timepoint, -->
<!--   #                    values_from = IgG, -->
<!--   #                    values_fn = list) %>% -->
<!--   ggplot(aes(x = End_to_sample,  -->
<!--              y = IgG,  -->
<!--              colour = factor(Dexa), -->
<!--              group = factor(PatID_Studie1))) + -->
<!--         labs(col = "Dexa") + -->
<!--         #theme(axis.text.x = element_text(angle = 90)) + -->
<!--         #geom_jitter(width = 0.1, alpha = 0.2) + -->
<!--         geom_point(position = position_dodge(0.4), alpha = 0.5) + -->
<!--         coord_trans(y = "log10") + -->
<!--         scale_y_continuous(breaks = c(10, 100, 1000, 10000)) + -->

<!--         #geom_boxplot(aes(fill = factor(Dexa)), width = 0.5,  -->
<!--         #             colour = "gray47", alpha = 0.5, outlier.shape = NA) + -->
<!--         scale_color_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values=c("orangered", "blue", "green3")) + -->
<!--         scale_fill_manual(breaks = c("D+", "D-", "mild"), -->
<!--                           values=c("orangered", "blue", "green3")) + -->
<!--         # facet_wrap(~factor(Marker2, levels = c("CD40L+ CD137+","CD40L+ GrzB+", -->
<!--         #                                        "CD40L+ IFN-g+","CD40L+ IL-21+", -->
<!--         #                                        "CD40L+ TNF-a+","CD137+ GrzB+", -->
<!--         #                                        "CD137+ IFN-g+","CD137+ IL-21+", -->
<!--         #                                        "CD137+ TNF-a+","CD137+ GrzB-")), -->
<!--         #            scales = "free") + -->
<!--         facet_wrap(. ~ factor(Dexa, levels = c('D+', 'D-', 'mild')), -->
<!--                  scales = "free") + -->
<!--                  #labeller = labeller(Dexa = as_labeller(dexa.labels))) + -->
<!--         geom_line(position = position_dodge(0.4),  -->
<!--                   aes(group = factor(PatID_Studie1)), alpha = 0.8,  -->
<!--                   size = 0.2) + -->
<!--         xlab("Time point") + -->
<!--         ggtitle("IgG evolution") + -->
<!--         theme_bw() + -->
<!--         #theme(legend.position = "none") + -->
<!--         theme(strip.background =element_rect(fill="aliceblue"))) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- ## How many patients have at least three time points in each measurement. -->
<!-- # facs_pat_3timepoints <- facs %>% -->
<!-- #   select(PatID_Studie1, Timepoint) %>% -->
<!-- #   distinct() %>% -->
<!-- #   group_by(PatID_Studie1) %>% -->
<!-- #   mutate(number_of_FACS_timepoints = n()) %>% -->
<!-- #   mutate(measured_at = paste0(Timepoint, collapse = ", ")) %>% -->
<!-- #   select(!Timepoint) %>% -->
<!-- #   distinct() -->
<!-- # write.table(facs_pat_3timepoints, "FACS_measurement_timepoints.tsv", row.names = F, quote = F, sep = "\t") -->
<!-- #  -->
<!-- # PNT_pat_3timepoints <- PNT %>% -->
<!-- #   select(PatID_Studie1, Days) %>% -->
<!-- #   distinct() %>% -->
<!-- #   group_by(PatID_Studie1) %>% -->
<!-- #   mutate(number_of_PNT_timepoints = n()) %>% -->
<!-- #   mutate(measured_at = paste0(Days, collapse = ", ")) %>% -->
<!-- #   select(!Days) %>% -->
<!-- #   distinct() -->
<!-- # write.table(PNT_pat_3timepoints, "PNT_measurement_timepoints.tsv", row.names = F, quote = F, sep = "\t") -->
<!-- #  -->
<!-- # sero_pat_3timepoints <- sero %>% -->
<!-- #   select(PatID_Studie1, Timepoint) %>% -->
<!-- #   distinct() %>% -->
<!-- #   group_by(PatID_Studie1) %>% -->
<!-- #   mutate(number_of_Sero_timepoint = n()) %>% -->
<!-- #   mutate(measured_at = paste0(Timepoint, collapse = ", ")) %>% -->
<!-- #   select(!Timepoint) %>% -->
<!-- #   distinct() -->
<!-- # write.table(sero_pat_3timepoints, "Serology_measurement_timepoints.tsv", row.names = F, quote = F, sep = "\t") -->

<!-- ``` -->

<!-- ```{r tables_for_Lara} -->
<!-- #  -->
<!-- # # convert facs table to wide format for selected markers, for each patient -->
<!-- # facs_wide <- read_excel("../20220104 Mastertable FACS incl. dead and VC.xlsx") %>%  -->
<!-- #   select("Studie 1",  -->
<!-- #          "PatID Studie 1",  -->
<!-- #          "timepoints 2022",  -->
<!-- #          "Dexa",  -->
<!-- #          "Marker 1",  -->
<!-- #          "Marker 2",  -->
<!-- #          "#S1-unstim log",  -->
<!-- #          "#S2-unstim log",  -->
<!-- #          "ausschliessen") %>% -->
<!-- #   mutate(ausschliessen = replace_na(ausschliessen, 0)) %>%  -->
<!-- #   rename(Studie1 = `Studie 1`,  -->
<!-- #          PatID_Studie1 = `PatID Studie 1`,  -->
<!-- #          Timepoint = `timepoints 2022`, -->
<!-- #          Marker1 = `Marker 1`,  -->
<!-- #          Marker2 = `Marker 2`, -->
<!-- #          S1 = `#S1-unstim log`,  -->
<!-- #          S2 = `#S2-unstim log`) %>% -->
<!-- #   filter(!is.na(S1) | !is.na(S2)) %>% -->
<!-- #   mutate(PatID_Studie1 = as.character(PatID_Studie1)) %>%  -->
<!-- #   mutate(Timepoint = replace(Timepoint, PatID_Studie1=="435", "W 2")) %>%  -->
<!-- #   filter(!(PatID_Studie1 == "183" |  -->
<!-- #            PatID_Studie1 == "22" |  -->
<!-- #            PatID_Studie1 == "53" |  -->
<!-- #            PatID_Studie1 == "88" |  -->
<!-- #            PatID_Studie1 == "115" |  -->
<!-- #            PatID_Studie1 == "120" |  -->
<!-- #            PatID_Studie1 == "133" |  -->
<!-- #            PatID_Studie1 == "153")) %>%  -->
<!-- #   filter(!(ausschliessen == 1)) %>%  -->
<!-- #   mutate(Studie1 = str_replace(Studie1, c("Pa-COVID-19"), "Pa-COVID")) %>%  -->
<!-- #   mutate(PatID_Studie1_short = paste(tolower(paste(substr(Studie1, 1, 2))),  -->
<!-- #                                tolower(paste(substr(Studie1, (nchar(Studie1)-1), nchar(Studie1)),  -->
<!-- #                                              PatID_Studie1,  -->
<!-- #                                              sep = "")), -->
<!-- #                                sep = "")) %>%  -->
<!-- #   mutate(Timepoint = str_replace(Timepoint, c("1"), "PI")) %>%  -->
<!-- #   mutate(Timepoint = str_replace(Timepoint, c("M2-4"), "M3")) %>%  -->
<!-- #   mutate(Timepoint = str_replace(Timepoint, c("after month 5"), "M6+")) %>%  -->
<!-- #   mutate(Timepoint = str_replace(Timepoint, c("W3-6"), "M1")) %>%  -->
<!-- #   mutate(Timepoint = str_replace(Timepoint, c("W 2"), "W2")) -->
<!-- #  -->
<!-- # ### S1 ### -->
<!-- #  -->
<!-- # # 1 # -->
<!-- # CD40CD137 <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ CD137+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S1) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S1, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S1", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40CD137$PI[[i]][1] -->
<!-- # CD40CD137$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40CD137$W2[[i]][1] -->
<!-- # CD40CD137$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40CD137$M1[[i]][1] -->
<!-- # CD40CD137$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40CD137$M3[[i]][1] -->
<!-- # CD40CD137$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40CD137$`M6+`[[i]][1] -->
<!-- # CD40CD137$`M6+` <-M5 -->
<!-- #  -->
<!-- # # 2 # -->
<!-- # CD40TNF <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ TNF-a+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S1) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S1, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S1", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40TNF$PI[[i]][1] -->
<!-- # CD40TNF$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40TNF$W2[[i]][1] -->
<!-- # CD40TNF$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40TNF$M1[[i]][1] -->
<!-- # CD40TNF$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40TNF$M3[[i]][1] -->
<!-- # CD40TNF$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40TNF$`M6+`[[i]][1] -->
<!-- # CD40TNF$`M6+` <-M5 -->
<!-- #  -->
<!-- #  -->
<!-- # # 3 # -->
<!-- # CD40IL21 <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ IL-21+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S1) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S1, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S1", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40IL21$PI[[i]][1] -->
<!-- # CD40IL21$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40IL21$W2[[i]][1] -->
<!-- # CD40IL21$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40IL21$M1[[i]][1] -->
<!-- # CD40IL21$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40IL21$M3[[i]][1] -->
<!-- # CD40IL21$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40IL21$`M6+`[[i]][1] -->
<!-- # CD40IL21$`M6+` <-M5 -->
<!-- #  -->
<!-- # # 4 # -->
<!-- # livecells <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "live cells") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S1) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S1, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S1", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$PI[[i]][1])) -->
<!-- #     postvac[i] <- livecells$PI[[i]][1] -->
<!-- # livecells$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$W2[[i]][1])) -->
<!-- #     W2[i] <- livecells$W2[[i]][1] -->
<!-- # livecells$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$M1[[i]][1])) -->
<!-- #     M1[i] <- livecells$M1[[i]][1] -->
<!-- # livecells$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$M3[[i]][1])) -->
<!-- #     M3[i] <- livecells$M3[[i]][1] -->
<!-- # livecells$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- livecells$`M6+`[[i]][1] -->
<!-- # livecells$`M6+` <-M5 -->
<!-- #  -->
<!-- # write.csv(CD40CD137, file = "../S1_CD40CD137.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40TNF, file = "../S1_CD40TNF.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40CD137, file = "../S1_CD40IL21.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40CD137, file = "../S1_livecells.csv", row.names = F, quote = F) -->
<!-- #  -->
<!-- #  -->

<!-- ``` -->

<!-- ```{r S2} -->
<!-- # ### S2 ### -->
<!-- #  -->
<!-- # # 1 # -->
<!-- # CD40CD137 <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ CD137+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S2) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S2, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S2", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40CD137$PI[[i]][1] -->
<!-- # CD40CD137$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40CD137$W2[[i]][1] -->
<!-- # CD40CD137$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40CD137$M1[[i]][1] -->
<!-- # CD40CD137$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40CD137$M3[[i]][1] -->
<!-- # CD40CD137$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40CD137)) -->
<!-- #   if(!is.null(CD40CD137$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40CD137$`M6+`[[i]][1] -->
<!-- # CD40CD137$`M6+` <-M5 -->
<!-- #  -->
<!-- # # 2 # -->
<!-- # CD40TNF <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ TNF-a+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S2) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S2, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S2", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40TNF$PI[[i]][1] -->
<!-- # CD40TNF$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40TNF$W2[[i]][1] -->
<!-- # CD40TNF$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40TNF$M1[[i]][1] -->
<!-- # CD40TNF$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40TNF$M3[[i]][1] -->
<!-- # CD40TNF$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40TNF)) -->
<!-- #   if(!is.null(CD40TNF$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40TNF$`M6+`[[i]][1] -->
<!-- # CD40TNF$`M6+` <-M5 -->
<!-- #  -->
<!-- #  -->
<!-- # # 3 # -->
<!-- # CD40IL21 <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "CD40L+ IL-21+") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S2) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S2, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S2", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$PI[[i]][1])) -->
<!-- #     postvac[i] <- CD40IL21$PI[[i]][1] -->
<!-- # CD40IL21$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$W2[[i]][1])) -->
<!-- #     W2[i] <- CD40IL21$W2[[i]][1] -->
<!-- # CD40IL21$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$M1[[i]][1])) -->
<!-- #     M1[i] <- CD40IL21$M1[[i]][1] -->
<!-- # CD40IL21$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$M3[[i]][1])) -->
<!-- #     M3[i] <- CD40IL21$M3[[i]][1] -->
<!-- # CD40IL21$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(CD40IL21)) -->
<!-- #   if(!is.null(CD40IL21$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- CD40IL21$`M6+`[[i]][1] -->
<!-- # CD40IL21$`M6+` <-M5 -->
<!-- #  -->
<!-- # # 4 # -->
<!-- # livecells <- facs_wide %>%  -->
<!-- #   filter(Marker2 == "live cells") %>%  -->
<!-- #   select(Studie1, -->
<!-- #          PatID_Studie1, -->
<!-- #          PatID_Studie1_short, -->
<!-- #          Dexa, -->
<!-- #          Timepoint, -->
<!-- #          Marker2, -->
<!-- #          S2) %>%  -->
<!-- #   group_by(Timepoint) %>% -->
<!-- #   mutate(row = row_number()) %>% -->
<!-- #   tidyr::pivot_wider(id_cols = c(PatID_Studie1, Studie1, PatID_Studie1_short, Dexa, Marker2),  -->
<!-- #                      names_from = Timepoint,  -->
<!-- #                      values_from = S2, -->
<!-- #                      values_fn = list) %>%  -->
<!-- #   mutate(Stimulation = rep("S2", nrow(.))) %>%  -->
<!-- #   mutate(PI = na_if(PI, c("NULL"))) %>% -->
<!-- #   mutate(W2 = na_if(W2, c("NULL"))) %>% -->
<!-- #   mutate(M1 = na_if(M1, c("NULL"))) %>% -->
<!-- #   mutate(M3 = na_if(M3, c("NULL"))) %>% -->
<!-- #   mutate(`M6+` = na_if(`M6+`, c("NULL"))) -->
<!-- #  -->
<!-- # postvac <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$PI[[i]][1])) -->
<!-- #     postvac[i] <- livecells$PI[[i]][1] -->
<!-- # livecells$PI <-postvac -->
<!-- #  -->
<!-- # W2 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$W2[[i]][1])) -->
<!-- #     W2[i] <- livecells$W2[[i]][1] -->
<!-- # livecells$W2 <-W2 -->
<!-- #  -->
<!-- # M1 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$M1[[i]][1])) -->
<!-- #     M1[i] <- livecells$M1[[i]][1] -->
<!-- # livecells$M1 <-M1 -->
<!-- #  -->
<!-- # M3 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$M3[[i]][1])) -->
<!-- #     M3[i] <- livecells$M3[[i]][1] -->
<!-- # livecells$M3 <-M3 -->
<!-- #  -->
<!-- # M5 <- c() -->
<!-- # for(i in 1:nrow(livecells)) -->
<!-- #   if(!is.null(livecells$`M6+`[[i]][1])) -->
<!-- #     M5[i] <- livecells$`M6+`[[i]][1] -->
<!-- # livecells$`M6+` <-M5 -->
<!-- #  -->
<!-- # write.csv(CD40CD137, file = "../S2_CD40CD137.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40TNF, file = "../S2_CD40TNF.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40CD137, file = "../S2_CD40IL21.csv", row.names = F, quote = F) -->
<!-- # write.csv(CD40CD137, file = "../S2_livecells.csv", row.names = F, quote = F) -->
<!-- ``` -->
